---
title: "Monitor model for traffic crashes"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---

```{r setup, include=FALSE}
library(flexdashboard)
library(dplyr)
library(ggplot2)
library(yardstick)
library(lubridate)
library(RSocrata)
library(jsonlite)
library(httr)
library(pins)
library(plotly)
library(leaflet)
library(reactable)

## get old metrics from RStudio Connect
board_register("rsconnect", server = "https://colorado.rstudio.com/rsc")
old_metrics <- pin_get("julia.silge/traffic-crash-metrics", board = "rsconnect")


## get new crashes from API
weeks_ago <- today() - weeks(2)
crash_url <- glue::glue("https://data.cityofchicago.org/Transportation/Traffic-Crashes-Crashes/85ca-t3if?$where=CRASH_DATE > '{weeks_ago}'")
crash_raw <- as_tibble(read.socrata(crash_url))
crash <- crash_raw %>%
  arrange(desc(crash_date)) %>%
  transmute(injuries = if_else(injuries_total > 0, "injuries", "none"),
            crash_date,
            crash_hour,
            report_type = if_else(report_type == "", "UNKNOWN", report_type),
            num_units,
            posted_speed_limit,
            weather_condition,
            lighting_condition,
            roadway_surface_cond,
            first_crash_type,
            trafficway_type,
            prim_contributory_cause,
            latitude, longitude) %>%
  na.omit()

## get predictions for new crashes from model
predictions <- POST("https://colorado.rstudio.com/rsc/traffic-crashes/predict", 
                    body = toJSON(select(crash, -injuries))) %>%
  content("text", encoding = "UTF-8") %>% 
  fromJSON()

## compute metrics for new crashes
metrics_by_week <- crash %>%
  bind_cols(.pred_injuries = predictions) %>%
  mutate(predicted = if_else(.pred_injuries > 0.5, "injuries", "none")) %>%
  mutate_if(is.character, as.factor) %>%
  mutate(crash_date = floor_date(as.Date(crash_date), unit = "week")) %>%
  group_by(crash_date) %>%
  metrics(injuries, predicted, .pred_injuries)

## compute total crashes
totals_by_week <- crash %>% 
  mutate(crash_date = floor_date(as.Date(crash_date), unit = "week")) %>% 
  count(crash_date, name = "crashes")

## combine old metrics and new metrics
new_metrics <- old_metrics %>%
  filter(!crash_date %in% metrics_by_week$crash_date) %>%
  bind_rows(metrics_by_week %>% left_join(totals_by_week)) %>%
  arrange(.metric, crash_date)

## pin back to RStudio Connect for next time
pin(new_metrics, "traffic-crash-metrics", board = "rsconnect")  
```


### Model metrics monitor how the bagged tree model performs over time compared to how it performed on training and testing data.

```{r}
## get training metrics expected from model
csv_url <- "https://raw.githubusercontent.com/juliasilge/modelops-playground/master/crash-api/crash-model-metrics.csv"
model_metrics <- readr::read_csv(csv_url)

p <- new_metrics %>%
  filter(.metric %in% c("accuracy", "roc_auc")) %>%
  ggplot(aes(crash_date, .estimate)) +
  geom_hline(aes(yintercept = .estimate, color = .metric), 
             data = model_metrics,
             alpha = 0.7, size = 1.2, lty = 2) +
  geom_line(aes(color = .metric)) +
  geom_point(aes(color = .metric, size = crashes), alpha = 0.8) +
  facet_wrap(~.metric, scales = "free_y", ncol = 1) +
  theme(legend.position = "none") +
  labs(x = NULL, y = NULL) +
  scale_size(range = c(0, 5)) +
  theme_light()

p <- ggplotly(p)
hide_legend(p)
```

*** 

This dashboard monitors a model for [traffic crashes in Chicago](https://data.cityofchicago.org/Transportation/Traffic-Crashes-Crashes/85ca-t3if) that predicts whether a crash involves an injury or not. Learn more about this model monitoring project here:

https://github.com/juliasilge/modelops-playground

The dashed lines correspond to the model metrics (accuracy and ROC AUC) measured on **testing** data when the model was trained.

### This dashboard updates every week, downloading new crash data from the Chicago API and computing model metrics.

```{r}
p2 <- crash %>%
  count(injuries) %>% 
  ggplot(aes(injuries, n)) + 
  geom_col() + 
  labs(x = NULL, y = "Number of crashes") +
  theme_minimal()

ggplotly(p2)
```

*** 

This plot shows the _new_ crashes used for monitoring since the last time this dashboard was generated.

This data exhibits class imbalance, which was addressed during modeling via downsampling.

### Which crashes were misclassified by the bagged tree model?

```{r}
crash %>%
  bind_cols(pred_injuries = predictions) %>%
  select(-crash_date, -latitude, -longitude) %>%
  filter((pred_injuries > 0.8 & injuries == "none") | (pred_injuries < 0.2 & injuries == "injuries")) %>%
  arrange(pred_injuries) %>%
  select(`probability of injury` = pred_injuries, everything()) %>%
  reactable(defaultSorted = c("probability of injury"), filterable = TRUE,
            defaultColDef = colDef(header = function(value) gsub("_", " ", value, fixed = TRUE)),
            columns = list(injuries = colDef(name = "any injuries?"))
  )
```

*** 

This table shows the _new_ crashes that our model did the worst job at classifying, in either direction.

Sort and filter to explore these misclassifications.


### How are the crashes distributed across Chicago?

```{r}
pal <- colorFactor(c("gray20", "palevioletred3"), domain = c("injuries", "none"))

crash %>%
  filter(latitude > 0) %>% 
  leaflet() %>% 
  addTiles() %>% 
  addCircleMarkers(color = ~pal(injuries)) %>%
  addLegend(title = "Crash type", pal = pal, values = ~injuries, opacity = 1)
```

***

[This dataset](https://data.cityofchicago.org/Transportation/Traffic-Crashes-Crashes/85ca-t3if) covers traffic crashes on city streets within Chicago city limits under the jurisdiction of the Chicago Police Department. 

This map, like the previous panels, shows the _new_ crashes used for monitoring since the last time this dashboard was generated.


